### 2021-10-26 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@5e43ee994e53d4600e33b8d78c7c7714955c5c2b 🚀

\hypertarget{is-this-decomposition-over-estimating-the-effect-of-rewiring}{%
\subsubsection{Is this decomposition over-estimating the effect of
``rewiring?''}\label{is-this-decomposition-over-estimating-the-effect-of-rewiring}}

@@

\hypertarget{numerical-experiment-the-decomposition-captures-the-roles-of-species-turnover-and-connectance-accurately}{%
\subsubsection{Numerical experiment: the decomposition captures the
roles of species turnover and connectance
accurately}\label{numerical-experiment-the-decomposition-captures-the-roles-of-species-turnover-and-connectance-accurately}}

Consider now two bipartite networks, which still have \(R\) species on
either side, but differ in their connectance (\(\rho_1\) and \(\rho_2\))
-- by maintaining the assumption that species on one side are shared
with probability \(p\), and that interactions between shared species are
rewired at probability \(q\), we can examine the effect of varying both
connectance and turnover on the value of the \(\beta\)-diversity
components. Note that, although not presented, we will drop the
multiplicative constant \(R^2\) from all calculations, as it is a common
factor for all values; again, this implies that the results presented
here are independant of network richness.

The number of unique links due to species turnover is

\[U = (1-p)(\rho_1 + \rho_2)\,,\]

which decreases with the proportion of shared species, but increases
with connectance. The number of links between shared species takes a
little more steps to calculate. First, amongst the \(pR^2\) species in
both sub-graphs, network 1 will have \(\rho_1 pR^2\), and network 2 will
have \(\rho_2 pR^2\). Because \(\rho_1 \neq \rho_2\), there are only
\(\text{min}(\rho_1, \rho_2)pR^2\) links that can be shared, a
proportion \(q\) of which will undergo re-wiring, and a proportion
\((1-q)\) of which will be shared. This leads to the expression (after
dropping \(R^2\)) for the number of shared links:

\[A = p (1-q) \text{min}(\rho_1, \rho_2)\,.\]

The number of unique links due to shared species is the sum of all links
in network 1 (\(\rho_1 R^2\)), minus the sum of the shared links
(\(AR^2\)) and the unique links due to species turnover
(\((1-p)\rho_1R^2\)); this same quantity is calculated in the same way
for the second networks, leading to (after dropping the multiplicative
constant \(R^2\) and some simplifications)

\[S = p (\rho_1 + \rho_2) - 2A\,.\]

Note that as expected, this last quantity scales with the proportion of
shared species (\(p\)) \emph{and} with connectance (as shared species
bring more of their interactions), but decreases with the size of the
shared links components. The consequences of varying \(\rho_2\) and
\(p\) are presented in fig.~\ref{fig:numexp3}.

\begin{figure}
\hypertarget{fig:commden}{%
\centering
\includegraphics{figures/common_denominator/components.png}
\caption{Effects of varying the connectance of the second network
(\(\rho_2\)) and the proportion of shared species (\(p\)) on the values
of the \(\beta\)-diversity components. As expected, \(\beta_{os}\) is
still independent of species turnover, and \(\beta_{wn}\) increases when
species turnover increases, or when the connectances become more
dissimilar. These figures have been generated with \(\rho_1 = 0.25\) and
\(q = 0.15\), and the results are qualitatively robust to changes in
these parameters.}\label{fig:commden}
}
\end{figure}

Although \(\beta_{os}\) is only responding to changes in connectance (as
is expected, seeing that the relative connectances of both networks
appear in the expression for \(S\) and \(A\)), \(\beta_{wn}\) changes in
response to both parameters. Specifically, increasing the difference in
connectance between the two networks, especially when also increasing
the species dissimilarity, results in more dissimilar networks -- this
is because unique species from both networks bring their own
interactions (at rate \(\rho_1\) and \(\rho_2\)), and therefore
contribute to dissimilarity. It is particularly noteworthy that
\(\beta_{st}\), regardless of the differences in connectance, increases
with the proportion of unique species. At an equal proportion of shared
species, \(\beta_{st}\) \emph{decreases} with differences in
connectance: this is an equally expected result, which indicates that
the difference between \(\beta_{os}\) and \(\beta_{wn}\) is in part
explained by non-turnover mechanisms (here, changes in connectance).
Relying on the \(\beta_{st}/\beta_{wn}\) correction again magnifies this
effect, without changing their interpretation.

@@

The choice of changing the denominator hinges on what one admits as a
definition for \(\beta_{st}\). If the point of \(\beta_{st}\) is to be a
component of overall \(\beta\)-diversity as advocated by Fründ (2021)
and Novotny (2009), a change of numerator \emph{might} be acceptable.
Nevertheless, this change of numerator contributes to blurring the
frontier between a measure of interaction dissimilarity and a measure of
community dissimilarity which starts to add the effect of relative
richness; this later case warrants a thorough methodological assessment.
Conversely, if as we argue in Poisot \emph{et al.} (2012),
\(\beta_{st}\) is to be meant as a \emph{guide} to the interpretation of
\(\beta_{wn}\) and \(\beta_{os}\), and related to actual measures of
species turnover and network connectance, one must not change the
denominator.

It is essential to recognize that there are multiple reasons to
calculate network dissimilarity, and it is our opinion that the
arguments levied by Fründ (2021) against the original partition stem
from a misunderstanding of what it intends to do (and does, indeed, do
well), not from intrinsic methodological issues in the partition itself.
Based on the results presented in this contribution, I argue that the
original partition of network \(\beta\)-diversity from Poisot \emph{et
al.} (2012) should remain the default.

@@

\leavevmode\hypertarget{ref-Novotny2009BetDiv}{}%
Novotny, V. (2009). Beta diversity of plant--insect food webs in
tropical forests: A conceptual framework. \emph{Insect Conservation and
Diversity}, 2, 5--9.

>>>>

@@

\hypertarget{using-an-alternative-normalization-trivializes-the-results}{%
\subsubsection{Using an alternative normalization trivializes the
results}\label{using-an-alternative-normalization-trivializes-the-results}}

In this numerical experiment, we reproduce the results in
fig.~\ref{fig:turnrew}, but using the alternative normalization
described above. The results are presented in fig.~\ref{fig:commden}.
Producing the analytical solutions for the various components, following
the expressions for \(C\), \(T\), and \(R\) given for
fig.~\ref{fig:turnrew}, yields a similar value for \(\beta_{wn}\)
(\emph{i.e.} the two approaches estimate the same value for total
dissimiliarity), but different values for \(\beta_{st}\) and
\(\beta_{os}\). Specifically, \(\beta_{os}\) becomes \(p^2(1-q)\), which
becomes dependent on species turnover. This, from an ecological point of
view, makes no sense: the quantification of how much shared species
interact in a similar way should not depend on how much species actually
overlap. The opposite problem arises for \(\beta_{st}\), which becomes
\(1-p^2\). In short, the relative importance of species turnover is
simply species turnover itself, and has no information on interaction
dissimilarity. Therefore the core issue of the Fründ (2021) alternative
is that, by attempting to fix a non-issue (namely the over-estimate of
the importance of re-wiring, which is only true in trivially small
networks), it blurs the meaning of \(\beta_{os}\), and renders
\(\beta_{st}\) useless as it is a re-expression of species
beta-diversity.

\begin{figure}
\hypertarget{fig:commden}{%
\centering
\includegraphics{figures/common_denominator/components.png}
\caption{Reproduction of fig.~\ref{fig:turnrew} with the alternative
denominators proposed by Fründ (2021).}\label{fig:commden}
}
\end{figure}

@@

Based on the numerical experiments and the derivations presented in this
paper, we can establish a number of recommendations for the measurement
and analysis of network dissimilarity. First, \(\beta_{os}\) allows to
estimate the rate of rewiring, which is an important ecological
information to have; quantifying it properly can give insights as to how
networks differ. Second, \(\beta_{st}\) captures both turnover and
rewiring mechanisms, but its interpretation is easier to accomplish in
the context of total network dissimilarity, and therefore
\(\beta_{st}/\beta_{wn}\) should be interpreted more thoroughly.
Finally, because the alternative denominator from Fründ (2021) removes
the interesting property of \(\beta_{os}\) (independent estimate of
rewiring rate), and trivializes the meaning of \(\beta_{st}\) (by
turning it into species dissimilarity), there seems to be no valid
reason to use it.

@@
